---
title: 'Assignment 1: Motor Trend report'
author: "C. Deramond"
date: "March, 2015"
output:
  html_document:
    highlight: pygments
    theme: flatly
---
# Executive summary

lalala

# Introduction

In this report we'll study the relationship between MPG and several features of vehicles. By the end of this study we hope to have answered these two questions:

1. *“Is an automatic or manual transmission better for MPG”*

2. *"Quantify the MPG difference between automatic and manual transmissions"*

All the data used in this report comes from `mtcars` dataset available on R datasets package.

# Analysis

## Exploratory

```{r, echo=FALSE, results='hide', cache=TRUE}
library(datasets)
library(ggplot2)
library(GGally)
tol_cor <- .55

yvx <- cor(mtcars)[1,]
selSet <- names(yvx[ which(abs(yvx) > tol_cor)]) # we will include only the ones with |cor| > .55
data<- mtcars[,selSet]

##grafico de pares
pairsXY <- ggpairs(
                data,
                title = "Y vs X, scatter matrix",
                upper = list(continuous = "cor"),
                lower = list(continuous = "smooth", 
                             params = c(method = "lm", fill = "red"))
)
##grafico de corr
corXY <- ggcorr(data) #+ ggtitle("Correlation matrix")
##

```

We'll start the process by discarding all variables which have a lower than `r tol_cor` correlation with our output `mpg`. This process lefts us with `r length(selSet) - 1` variables. Time to look for multicolinearity:

If you see [Graph 01](#g01), the risk for muticolinearity (i.e. our predictor are not independent among each other) is high.


## Regression


### Goodness of fit


# Results



# Appendix


<a name="g01"></a>**Graph 1: Correlation matrix**

```{r, echo=FALSE}
print(corXY)

```

<a name="g02"></a>**Graph 2: Scatter matrix, all variables**

```{r, echo=FALSE, cache=TRUE}
print(pairsXY)

```



