---
title: 'Assignment 1: Motor Trend report'
author: "C. Deramond"
date: "March, 2015"
output:
  pdf_document: default
  html_document:
    highlight: pygments
    theme: flatly
---
# Executive summary

In this document we explored car performance in terms of miles per gallon (mpg). By methods of regression, we conclude that manual transmission will give better performance in a boost of app 1.8 MPG, ceteris paribus.

Alas the question are answered by the model in a pretty straightforward way. In the following sections you may review the procedure as well as the validations of the model.


# Introduction

In this report we'll study the relationship between MPG and several features of vehicles. By the end of this study we hope to have answered these two questions:

1. *“Is an automatic or manual transmission better for MPG”*

2. *"Quantify the MPG difference between automatic and manual transmissions"*

All the data used in this report comes from `mtcars` dataset available on R datasets package.

# Analysis

## Exploratory

```{r chunkLoad, echo=FALSE, results='hide', cache=TRUE}
library(datasets)
library(ggplot2)
library(GGally)
library(car)
tol_cor <- .55

yvx <- cor(mtcars)[1,]
selSet <- names(yvx[ which(abs(yvx) > tol_cor)]) # we will include only the ones with |cor| > .55
data<- mtcars[,selSet]
#data$PWratio <- data$hp/data$wt

##grafico de pares
pairsXY <- ggpairs(
                data,
                title = "Y vs X, scatter matrix",
                upper = list(continuous = "cor"),
                lower = list(continuous = "smooth", 
                             params = c(method = "lm", fill = "red"))
)
##grafico de corr
corXY <- ggcorr(data) #+ ggtitle("Correlation matrix")
##

```

We'll start the process by discarding all variables which have a lower than `r tol_cor` correlation with our output `mpg`. This process lefts us with `r length(selSet) - 1` variables.

A quick [scatter matrix](#g01) shows many interesting relationships that we should address:

- ` cyl, vs, am ` are discrete variable variables. We take out `vs` as it is of low interest. And turn the rest into factor
```{r chunkTransforms, echo=FALSE}
data<- data[,-7]
data$am <- factor(data$am, labels = c("Auto", "Manual"))
data$cyl <- factor(data$cyl)


```

        
- the correlation within the predictors is quite high, so risk of multicollinearity is there.

If you see [Graph 02](#g02), the risk for muticolinearity (i.e. our predictors are not independent among each other) is high.


## Regression

```{r chunkRegs, echo=FALSE, results='hide'}
fit_full <- lm(mpg ~ ., data = data)
fit_sw <- step(fit_full)

#cyl +hp+wt+am
fit1 <- update(fit_full, formula. = mpg ~ am)
fit2 <- update(fit_full, formula. = mpg ~ am + cyl )
fit3 <- update(fit_full, formula. = mpg ~ am + cyl + hp)
fit4 <- update(fit_full, formula. = mpg ~ am + cyl + hp + wt + interaction(cyl, am))

```

As starting point, let's consider the suggested model by stepwise: **`r as.character(fit_sw$call[2])`**  

As we are including 2 factors, interaction will be considered [see interaction plot](#g03). A second look at the ANOVA table, tells that interaction is not significant:

```{r, echo=FALSE}
anova(fit4)
fit4 <- update(fit4, .~.-interaction(cyl, am))
#fit4_inter <- update(fit_full, formula. = mpg ~ cyl + hp + wt + am + interaction(cyl, am))

```

We rerun model without interaction, thus suggesting using manual transmission has a gain of `r round(fit4$coefficients[2],2)` in mpg, which makes sense. Time to evaluate our model.

```{r echo=FALSE}
summary (fit4)
```


### Model evaluation

<h4> Nested ANOVA </h4>

as we have merely `r length(fit4$model)-1` predictor variables, a nested ANOVA, suffices to prove our variables are indeed adding explicative value. As it shows below, 

```{r, echo=FALSE}
anova(fit1, fit2, fit3, fit4)

```

<h4> Model diagnostics & validation </h4>

Graphically we can examine our normality and homoskedacity assumptions [see graphs 4](#g04). Additionally a validation of components vs residuals can be seen on [graphs 5](#g05)



# Results

The model as presented by its coefficients is self explanatory:

```{r, echo=FALSE}
fit4$coefficients

```

Manual transmission cars will put out better mileage app `r round(fit4$coefficients[2],2)`, ceteris paribus. 


# Appendix


<a name="g01"></a>**Graph 1: Scatter matrix, all variables**

```{r, echo=FALSE, cache=TRUE, dependson="chunkLoad"}
print(pairsXY)

```

<a name="g02"></a>**Graph 2: Correlation matrix**

```{r, echo=FALSE, dependson="chunkLoad"}
print(corXY)

```

<a name="g03"></a>**Graph 3: Interaction plot am vs  cyl**

```{r, echo=FALSE}
interaction.plot(data$cyl, data$am, response = data$mpg)

```

<a name="g04"></a>**Graph 4: Diagnostics plots for model**

```{r, echo=FALSE}
# par(mfcol = c(2,2))
plot(fit4)

```

<a name="g05"></a>**Graph 5: Components and Residual plots**

```{r, echo=FALSE,  dependson="chunkLoad"}
library(car)
crPlots(model = fit4)

```

